---
title: "Untitled"
output: html_document
---

```{r}
regular.df <- read.csv("./womens-machine-learning-competition-2019/WDataFiles/WRegularSeasonDetailedResults.csv")
str(regular.df)

tourney.df2 <- readRDS("tourney.df2.rds") 
str(tourney.df2)
```

Definition of Adjusted Win Percentage:

Home Wins and Away Losses are weighted at 0.6.

Neutral results are weighted at 1.

Home Losses and Away Wins are weighted at 1.4.

Eg Adjusted Win Percentage of:

2 Home Wins

1 Away Loss

1 Home Loss

= (2x0.6) / (2x0.6 + 1x0.6 + 1x1.4)

= 37.5% (but their Win Percentage is 50%)

Let's represent this definition as a function for a given set of matches.

```{r}
calculate.adj.win.percentage <- function(teamID, regular.df){
  
  #neutral wins and losses
  neutral.matches.df <- regular.df[as.character(regular.df$WLoc) == "N",]
  neutral.wins <- sum(neutral.matches.df$WTeamID == teamID)
  neutral.losses <- sum(neutral.matches.df$LTeamID == teamID)
  
  #home and away wins
  win.matches <- regular.df[regular.df$WTeamID == teamID,]
  away.wins <- sum(as.character(win.matches$WLoc) == "A")
  home.wins <- sum(as.character(win.matches$WLoc) == "H")
  
  #home and away losses
  loss.matches <- regular.df[regular.df$LTeamID == teamID,]
  home.losses <- sum(as.character(loss.matches$WLoc) == "A")
  away.losses <- sum(as.character(loss.matches$WLoc) == "H")
  
  #weight the values up
  away.wins <- away.wins*1.4
  home.wins <- home.wins*0.6
  home.losses <- home.losses*1.4
  away.losses <- away.losses*0.6
  
  adj.win.pctg <- (neutral.wins+away.wins+home.wins)/(neutral.wins+away.wins+home.wins+neutral.losses+away.losses+home.losses)
  return(adj.win.pctg)
}
```

STEP ONE:

Calculate Adjusted Win Percentage for each team against March Madness teams (if NaN, record as 0%, this is an ASSUMPTION)

```{r}
wteam.round1.adj.win.percentage <- c()
lteam.round1.adj.win.percentage <- c()
for(i in 1:nrow(tourney.df2)){
  
  tournament.match <- tourney.df2[i,]
  tournament.match.year <- tournament.match$Season
  tournament.match.wteam <- tournament.match$WTeamID
  tournament.match.lteam <- tournament.match$LTeamID
  
  #subset regular.df into the year of tournament match
  regular.df.year <- regular.df[regular.df$Season == tournament.match.year,]
  
  #subset regular.df.year into matches between march madness teams

  tourney.df2.year <- tourney.df2[tourney.df2$Season == tournament.match.year,]

  march.madness.teams <- unique(c(tourney.df2.year$WTeamID,tourney.df2.year$LTeamID))

  regular.df.year.march.teams <- regular.df.year[which(c(regular.df.year$WTeamID %in% march.madness.teams) & c(regular.df.year$LTeamID %in% march.madness.teams)),]
  
  #calculate adj.win.wteam
  wteam.regular.df.year.march.teams <- regular.df.year.march.teams[c(c(regular.df.year.march.teams$WTeamID == tournament.match.wteam) | c(regular.df.year.march.teams$LTeamID == tournament.match.wteam)),]

  adj.win.wteam <- calculate.adj.win.percentage(tournament.match.wteam, wteam.regular.df.year.march.teams)
  wteam.round1.adj.win.percentage <- c(wteam.round1.adj.win.percentage, adj.win.wteam)
  
  #calculate adj.win.lteam
  lteam.regular.df.year.march.teams <- regular.df.year.march.teams[c(regular.df.year.march.teams$WTeamID == tournament.match.lteam) | c(regular.df.year.march.teams$LTeamID == tournament.match.lteam),]

  adj.win.lteam <- calculate.adj.win.percentage(tournament.match.lteam, lteam.regular.df.year.march.teams)
  lteam.round1.adj.win.percentage <- c(lteam.round1.adj.win.percentage, adj.win.lteam)
}

unique(wteam.round1.adj.win.percentage)
unique(lteam.round1.adj.win.percentage)

lteam.round1.adj.win.percentage <- replace(lteam.round1.adj.win.percentage, which(is.nan(lteam.round1.adj.win.percentage)), 0)

unique(lteam.round1.adj.win.percentage)
```

STEP TWO:

Calculate Adjusted Win Percentage for each team against teams seeded 1 to 8 (if NaN, record as 0%, this is an ASSUMPTION)

```{r}
wteam.round2.adj.win.percentage <- c()
lteam.round2.adj.win.percentage <- c()

for(i in 1:nrow(tourney.df2)){
  
  tournament.match <- tourney.df2[i,]
  tournament.match.year <- tournament.match$Season
  tournament.match.wteam <- tournament.match$WTeamID
  tournament.match.lteam <- tournament.match$LTeamID
  
  #subset regular.df into the year of tournament match
  regular.df.year <- regular.df[regular.df$Season == tournament.match.year,]
  
  #subset regular.df.year into matches between march madness teams seeded 1 to 8
  tourney.df2.year <- tourney.df2[tourney.df2$Season == tournament.match.year,]
  
  wteamID.of.1to8seeds <- tourney.df2.year$WTeamID[tourney.df2.year$wteam.seeds <= 8]
  lteamID.of.1to8seeds <- tourney.df2.year$LTeamID[tourney.df2.year$lteam.seeds <= 8]
  teams.1to8seeds <- unique(c(wteamID.of.1to8seeds, lteamID.of.1to8seeds))
  
  regular.df.year.1to8seeds <- regular.df.year[which(c(regular.df.year$WTeamID %in% teams.1to8seeds) & c(regular.df.year$LTeamID %in% teams.1to8seeds)),]
  
  #calculate adj.win.wteam
  wteam.regular.df.year.1to8seeds <- regular.df.year.1to8seeds[c(c(regular.df.year.1to8seeds$WTeamID == tournament.match.wteam) | c(regular.df.year.1to8seeds$LTeamID == tournament.match.wteam)),]

  adj.win.wteam <- calculate.adj.win.percentage(tournament.match.wteam, wteam.regular.df.year.march.teams)
  wteam.round2.adj.win.percentage <- c(wteam.round2.adj.win.percentage, adj.win.wteam)
  
  #calculate adj.win.lteam
  lteam.regular.df.year.1to8seeds <- regular.df.year.1to8seeds[c(c(regular.df.year.1to8seeds$WTeamID == tournament.match.lteam) | c(regular.df.year.1to8seeds$LTeamID == tournament.match.lteam)),]

  adj.win.lteam <- calculate.adj.win.percentage(tournament.match.lteam, lteam.regular.df.year.march.teams)
  lteam.round2.adj.win.percentage <- c(lteam.round2.adj.win.percentage, adj.win.lteam)
  
}

unique(wteam.round2.adj.win.percentage)
table(wteam.round2.adj.win.percentage)
unique(lteam.round2.adj.win.percentage)
table(lteam.round2.adj.win.percentage)
```

We notice the AdjWin% metric against teams seeded from 1 to 8 has too much missing data. Therefore, use AdjWin% metric against teams that reached March Madness for the model.

```{r}
tourney.df3 <- cbind.data.frame(tourney.df2, wteam.round1.adj.win.percentage, lteam.round1.adj.win.percentage)
str(tourney.df3)
saveRDS(tourney.df3, "tourney.df3.rds")
```