---
title: "Untitled"
output: html_document
---

Copy, Paste and Tweak appropriate code from Mens comp

```{r}
regular.df <- read.csv("./womens-machine-learning-competition-2019/WDataFiles/WRegularSeasonDetailedResults.csv")
str(regular.df)

tourney.df3 <- readRDS("tourney.df3.rds")
str(tourney.df3)
```

```{r}
#Function: for a given year and teamID, return the relevant INDEXES and winning team y/n vector on regular.df

year.teamID.into.indexes <- function(year, teamID, regular.df){
  
  indexes <- which(regular.df$Season == year & regular.df$WTeamID == teamID | regular.df$Season == year & regular.df$LTeamID == teamID)
  
  wins <- which(regular.df$Season == year & regular.df$WTeamID == teamID)
  
  winner.logical <- indexes %in% wins
  
  return(cbind(indexes, winner.logical))
}
```

Net Efficiency definition:

Net Efficiency = Offensive Efficiency - Defensive Efficiency

Offensive Efficiency = Total Points/Total Possesion

Total Possesion = Field Goal Attempts - Offensive Rebounds + Turnovers + .475*Free Throw Attempts

Same logic for defensive efficiency

WFGA is field goal attempts by winning team

WOR is offensive rebounds by winning team

WTO is turnovers COMMITTED (Google meaning of this) by winning team

WFTA is free throw attemps by winning team

```{r}
#Function: for a given set of rows (using the INDEXES and winning team y/n) on regular.df, return the average Net Efficiency

indexes.for.regular.into.neteff <- function(indexes, winner.logical, regular.df, year, tourney.df3){
  
  relevant.regular.matches.df <- regular.df[indexes,]
  
  #find teamID of all teams that reached this particular year's march madness
  tourney.df3.year <- tourney.df3[tourney.df3$Season == year,]
  march.madness.teams <- unique(c(tourney.df3.year$WTeamID, tourney.df3.year$LTeamID))
  
  #subset relevant.regular.matches.df into matches against teams that reached march madness
  march.madness.indexes <- which(c(relevant.regular.matches.df$WTeamID %in% march.madness.teams) & c(relevant.regular.matches.df$LTeamID %in% march.madness.teams))
  winner.logical <- winner.logical[march.madness.indexes]
  relevant.regular.matches.df <- relevant.regular.matches.df[march.madness.indexes,]
  
  #Compute offensive efficiency
  fga <- c()
  offr <- c()
  turno <- c()
  fta <- c()
  score <- c()
  
  if(length(march.madness.indexes) == 0){#if a tam didin't play any March Madness teams during regular season
    return(-1)
  }else{
    for(i in 1:nrow(relevant.regular.matches.df)){
      regular.match <- relevant.regular.matches.df[i,]
      if(winner.logical[i]){#if winner of regular.match
        fga <- c(fga, regular.match$WFGA)
        offr <- c(offr, regular.match$WOR)
        turno <- c(turno, regular.match$WTO)
        fta <- c(fta, regular.match$WFTA)
        score <- c(score, regular.match$WScore)
      }else{#if loser of regular.match
        fga <- c(fga, regular.match$LFGA)
        offr <- c(offr, regular.match$LOR)
        turno <- c(turno, regular.match$LTO)
        fta <- c(fta, regular.match$LFTA)
        score <- c(score, regular.match$LScore)
      }
    }
    
    fga <- mean(fga)
    offr <- mean(offr)
    turno <- mean(turno)
    fta <- mean(fta)
    score <- mean(score)
    
    total.posession <- fga - offr + turno + (fta*.475)
    off.eff <- score / total.posession
    
    #Compute defensive efficiency
    fga <- c()
    offr <- c()
    turno <- c()
    fta <- c()
    score <- c()
    
    for(i in 1:nrow(relevant.regular.matches.df)){
      regular.match <- relevant.regular.matches.df[i,]
      if(winner.logical[i]){#if winner of regular.match
        fga <- c(fga, regular.match$LFGA)
        offr <- c(offr, regular.match$LOR)
        turno <- c(turno, regular.match$LTO)
        fta <- c(fta, regular.match$LFTA)
        score <- c(score, regular.match$LScore)
      }else{#if loser of regular.match
        fga <- c(fga, regular.match$WFGA)
        offr <- c(offr, regular.match$WOR)
        turno <- c(turno, regular.match$WTO)
        fta <- c(fta, regular.match$WFTA)
        score <- c(score, regular.match$WScore)
      }
    }
    
    fga <- mean(fga)
    offr <- mean(offr)
    turno <- mean(turno)
    fta <- mean(fta)
    score <- mean(score)
    
    def.possesion <- fga - offr + turno + (fta*.475)
    def.eff <- score / def.possesion
    
    #Compute net efficiency
    net.eff <- off.eff - def.eff
    return(net.eff)
  }
}
```

```{r}
#Create wteam.efficiency and lteam.efficiency using the TeamID (W and L TeamID) and Season data from tourney.df to find the appropriate indexes to insert each average Net Efficiency computed

wteam.efficiency <- c()
lteam.efficiency <- c()

for(i in 1:nrow(tourney.df3)){
  tourney.match <- tourney.df3[i,]
  year <- tourney.match$Season
  team1ID <- tourney.match$WTeamID
  team2ID <- tourney.match$LTeamID
  
  team1.indexes.logical <- year.teamID.into.indexes(year, team1ID, regular.df)
  team2.indexes.logical <- year.teamID.into.indexes(year, team2ID, regular.df)
  
  team1.net.eff <- indexes.for.regular.into.neteff(team1.indexes.logical[,1], team1.indexes.logical[,2], regular.df, year, tourney.df3)
  team2.net.eff <- indexes.for.regular.into.neteff(team2.indexes.logical[,1], team2.indexes.logical[,2], regular.df, year, tourney.df3)
  
  wteam.efficiency <- c(wteam.efficiency, team1.net.eff)
  lteam.efficiency <- c(lteam.efficiency, team2.net.eff)
}

length(unique(lteam.efficiency))
nrow(tourney.df3)
#should be same lengths because when a team with a certain net efficiency loses, that team will never play again.
```

```{r}
tourney.df4 <- cbind.data.frame(tourney.df3, wteam.efficiency, lteam.efficiency)
str(tourney.df4)
saveRDS(tourney.df4, "tourney.df4.rds")
```