---
title: "Untitled"
output: html_document
---

```{r}
tourney.df4 <- readRDS("tourney.df4.rds")
str(tourney.df4)
```

Change WLoc into 1 for Home, 0 for Neutral and -1 for Away

```{r}
location <- as.integer(tourney.df4$WLoc)
location[location == 1] <- -1
location[location == 2] <- 1
location[location == 3] <- 0

tourney.df4$WLoc <- location
str(tourney.df4)
```

To compute the form metric, mainly Copy, Paste and Tweak appropriate code from Mens comp.

```{r}
#Function: for a given seed (such as "W01"), return the sequence of the strongest seeds a team can clash against during its road to the championship crown

opponent.seed.sequence <- function(seed){
  
  seed.sequence <- c(17-seed) #round 1 is obvious
  if(seed.sequence >= 10){
    seed.sequence <- as.character(seed.sequence)
  }else{
    seed.sequence <- paste0("0", seed.sequence)
  }
  
  ncaa.brackets <- c(1, 16, 8, 9, 5, 12, 4, 13, 6, 11, 3, 14, 7, 10, 2, 15)
  seed.pos <- which(ncaa.brackets == seed)
  #ROUND 2 START##########################
  if(ceiling(seed.pos / 4) == 1){ #first quarter of 16
    if(seed != 1){
      seed.sequence <- c(seed.sequence, "01")
    }else{
      seed.sequence <- c(seed.sequence, "08")
    }
  }
  else if(ceiling(seed.pos / 4) == 2){ #second quarter of 16
    if(seed != 4){
      seed.sequence <- c(seed.sequence, "04")
    }else{
      seed.sequence <- c(seed.sequence, "05")
    }
  }
  else if(ceiling(seed.pos / 4) == 3){ #third quarter of 16
    if(seed != 3){
      seed.sequence <- c(seed.sequence, "03")
    }else{
      seed.sequence <- c(seed.sequence, "06")
    }
  }
  else{#fourth quarter of 16
    if(seed != 2){
      seed.sequence <- c(seed.sequence, "02")
    }else{
      seed.sequence <- c(seed.sequence, "07")
    }
  }
  #ROUND 2 END##########################
  #ROUND 3 START##########################
  if(ceiling(seed.pos / 8) == 1){ #first half of 16
    if(seed != 1){
      seed.sequence <- c(seed.sequence, "01")
    }else{
      seed.sequence <- c(seed.sequence, "04")
    }
  }
  else{#second half of 16
    if(seed != 2){
      seed.sequence <- c(seed.sequence, "02")
    }else{
      seed.sequence <- c(seed.sequence, "03")
    }
  }
  #ROUND 3 END##########################
  #ROUND 4, 5 START##########################
  if(seed != 1){
    seed.sequence <- c(seed.sequence, "01", "01", "01")
  }else{
    seed.sequence <- c(seed.sequence, "02", "01", "01")
  }
  #ROUND 4, 5 END##########################
  
  return(as.integer(seed.sequence)) #TWEAK to the copy and pasted code
  
  #ROUND 6 is not required because no match will have this round as the previous round
}
```

Use the function above to return the strongest possible opponent wteam and lteam beat in the previous round.

The function above returns the sequence of strongest possible opponents.

```{r}
#Create round.they.clash

round.they.clash <- c()

for(i in 1:nrow(tourney.df4)){
  
  row <- tourney.df4[i,]
  day <- row$DayNum
  
  if(row$Season >= 2017){
    if(day <= 138){
      round.they.clash <- c(round.they.clash, 1) #round of 68 counts as round1
    }else if(day <= 140){
      round.they.clash <- c(round.they.clash, 2)
    }else if(day <= 145){
      round.they.clash <- c(round.they.clash, 3)
    }else if(day <= 147){
      round.they.clash <- c(round.they.clash, 4)
    }else if(day == 151){
      round.they.clash <- c(round.they.clash, 5)
    }else{
      round.they.clash <- c(round.they.clash, 6)
    }
  }else if (row$Season >= 2015){
      if(day <= 138){
        round.they.clash <- c(round.they.clash, 1) #round of 68 counts as round1
      }else if(day <= 140){
        round.they.clash <- c(round.they.clash, 2)
      }else if(day <= 145){
        round.they.clash <- c(round.they.clash, 3)
      }else if(day <= 147){
        round.they.clash <- c(round.they.clash, 4)
      }else if(day == 153){
        round.they.clash <- c(round.they.clash, 5)
      }else{
        round.they.clash <- c(round.they.clash, 6)
      }
  }else{
      if(day <= 139){
        round.they.clash <- c(round.they.clash, 1) #round of 68 counts as round1
      }else if(day <= 141){
        round.they.clash <- c(round.they.clash, 2)
      }else if(day <= 146){
        round.they.clash <- c(round.they.clash, 3)
      }else if(day <= 148){
        round.they.clash <- c(round.they.clash, 4)
      }else if(day == 153){
        round.they.clash <- c(round.they.clash, 5)
      }else{
        round.they.clash <- c(round.they.clash, 6)
      }
  }
  
}
```



```{r}
wteam.forms <- c()
lteam.forms <- c()

for(i in 1:nrow(tourney.df4)){
  row <- tourney.df4[i,]
  round <- round.they.clash[i]
  
  if(round == 1){
    wteam.forms <- c(wteam.forms, 16)
    lteam.forms <- c(lteam.forms, 16)
  }else{
    wteam.seed <- tourney.df4$wteam.seeds[i]
    lteam.seed <- tourney.df4$lteam.seeds[i]
    
    wteam.form <- opponent.seed.sequence(wteam.seed)[round-1]
    lteam.form <- opponent.seed.sequence(lteam.seed)[round-1]
    
    wteam.forms <- c(wteam.forms, wteam.form)
    lteam.forms <- c(lteam.forms, lteam.form)
  }
}
```

```{r}
tourney.df5 <- cbind.data.frame(tourney.df4, wteam.forms, lteam.forms)
str(tourney.df5)
saveRDS(tourney.df5, "tourney.df5.rds")
```

```{r}
tourney.df5 <- readRDS("tourney.df5.rds")
```

```{r}
#Home Advantage metric
home.advantage <- tourney.df5$WLoc

#Seed metric
seed.difference <- tourney.df5$wteam.seeds - tourney.df5$lteam.seeds
seed.difference.log <- log(seed.difference+16)
seed.ratio <- tourney.df5$wteam.seeds / tourney.df5$lteam.seeds
seed.ratio.log <- log(seed.ratio+1)

#Form metric
form.difference <- tourney.df5$wteam.forms - tourney.df5$lteam.forms
form.difference.log <- log(form.difference+16)
form.ratio <- tourney.df5$wteam.forms / tourney.df5$lteam.forms
form.ratio.log <- log(form.ratio+1)

#AdjustedWinP metric
adjwinp.difference <- tourney.df5$wteam.round1.adj.win.percentage - tourney.df5$lteam.round1.adj.win.percentage
adjwinp.difference.log <- log(adjwinp.difference+1)
adjwinp.ratio <- tourney.df5$wteam.round1.adj.win.percentage / tourney.df5$lteam.round1.adj.win.percentage
adjwinp.ratio.log <- log(adjwinp.ratio+1)

#Efficiency metric
efficiency.difference <- tourney.df5$wteam.efficiency - tourney.df5$lteam.efficiency
efficiency.difference.log <- log(efficiency.difference+2)
efficiency.ratio <- tourney.df5$wteam.efficiency / tourney.df5$lteam.efficiency
efficiency.ratio.log <- log(efficiency.ratio+2)
```

```{r}
#bind all metrics into one df
metrics.df <- cbind.data.frame(
  home.advantage, seed.difference, seed.difference.log, seed.ratio, seed.ratio.log, form.difference, form.difference.log, form.ratio, form.ratio.log, adjwinp.difference, adjwinp.difference.log, adjwinp.ratio, adjwinp.ratio.log, efficiency.difference, efficiency.difference.log, efficiency.ratio, efficiency.ratio.log
)
str(metrics.df)

#metrics vector
results <- c(rep(T, nrow(metrics.df)), rep(F, nrow(metrics.df)))
```

Create a dummy metrics.df and then merge so glm can be performed.

```{r}
#Home Advantage metric
dummy.home.advantage <- tourney.df5$WLoc*-1

#Seed metric
dummy.seed.difference <- tourney.df5$lteam.seeds - tourney.df5$wteam.seeds
dummy.seed.difference.log <- log(dummy.seed.difference+16)
dummy.seed.ratio <- tourney.df5$lteam.seeds / tourney.df5$wteam.seeds
dummy.seed.ratio.log <- log(dummy.seed.ratio+1)

#Form metric
dummy.form.difference <- tourney.df5$lteam.forms - tourney.df5$wteam.forms
dummy.form.difference.log <- log(dummy.form.difference+16)
dummy.form.ratio <- tourney.df5$lteam.forms / tourney.df5$wteam.forms
dummy.form.ratio.log <- log(dummy.form.ratio+1)

#AdjustedWinP metric
dummy.adjwinp.difference <- tourney.df5$lteam.round1.adj.win.percentage - tourney.df5$wteam.round1.adj.win.percentage
dummy.adjwinp.difference.log <- log(dummy.adjwinp.difference+1)
dummy.adjwinp.ratio <- tourney.df5$lteam.round1.adj.win.percentage / tourney.df5$wteam.round1.adj.win.percentage
dummy.adjwinp.ratio.log <- log(dummy.adjwinp.ratio+1)

#Efficiency metric
dummy.efficiency.difference <- tourney.df5$lteam.efficiency - tourney.df5$wteam.efficiency
dummy.efficiency.difference.log <- log(dummy.efficiency.difference+2)
dummy.efficiency.ratio <- tourney.df5$lteam.efficiency / tourney.df5$wteam.efficiency
dummy.efficiency.ratio.log <- log(dummy.efficiency.ratio+2)
```

```{r}
#bind all metrics into one df
dummy.metrics.df <- cbind.data.frame(
  dummy.home.advantage, dummy.seed.difference, dummy.seed.difference.log, dummy.seed.ratio, dummy.seed.ratio.log, dummy.form.difference, dummy.form.difference.log, dummy.form.ratio, dummy.form.ratio.log, dummy.adjwinp.difference, dummy.adjwinp.difference.log, dummy.adjwinp.ratio, dummy.adjwinp.ratio.log, dummy.efficiency.difference, dummy.efficiency.difference.log, dummy.efficiency.ratio, dummy.efficiency.ratio.log
)
colnames(dummy.metrics.df) <- c(
  "home.advantage", "seed.difference", "seed.difference.log", "seed.ratio", "seed.ratio.log", "form.difference", "form.difference.log", "form.ratio", "form.ratio.log", "adjwinp.difference", "adjwinp.difference.log", "adjwinp.ratio", "adjwinp.ratio.log", "efficiency.difference", "efficiency.difference.log", "efficiency.ratio", "efficiency.ratio.log"
)
str(dummy.metrics.df)
```

Merge the two dataframes and the results vector together

```{r}
metrics.df <- rbind.data.frame(metrics.df, dummy.metrics.df)
results.metrics.df <- cbind.data.frame(results, metrics.df)
saveRDS(results.metrics.df, "results.metrics.df.rds")
```
